### 2023-04-11 refereeResponse.txt
Edits up to fig 10

cript at hand presents a numerical study of planetesimal accretion. The authors identify a novel mode of accretion that operates if the primary's radius is a significant fraction of its hill-radius - a criterion the authors find to be met at short orbital periods in circumstellar disks. This new accretion mode is shown to be highly efficient, resulting in several large planetary embryos and a depletion of the primordial planetesimal formation, and is therefore distinct from bimodal planetesimal distribution that is the end product of an oligarchic growth phase.

I find the presented research to be innovative, interesting, and valuable for the scientific community. However, before I can recommend the manuscript for publication, I believe there are several areas that require improvement. Specifically, I suggest that the authors address the five major comments outlined below to ensure that the findings are properly supported and effectively communicated. I also include a detailed list of specific notes, comments and questions that I suggest the authors consider when revising the manuscript. I appreciate the authors' work and look forward to seeing the progress on the manuscript.

Major Comments

1) I am unconvinced that the orbital period of P = 60 days is such a clear-cut, and universal boundary between oligarchic accretion mode and the new 'high-alpha' accretion mode. As far as I understand, the authors motivate this value mainly from their fdHi simulation e.g. Fig. 5, when the value for tilde b of the largest bodies falls below the critical value of 2sqrt(3). I would naively argue that for fdHiShallow and for fdLow, this is happening at ~40 and ~70 days respectively (Fig. 8). A key feature of the oligarchic growth mode seems to be the production of a bimodal distribution of bodies. Even for fdHi, this is the case up to an inner orbit of P = 10 d. Next, in Fig.6 which plots the evolution of the primordial planetesimal surface density, I do not see a clear qualitatively different behavior between the curves for 50 and 60 orbits, vs the curves for 60 or 70 orbits. Yes, the depletion of planetesimals becomes more rapid the more one moves towards the star, but to first order, I would imagine this is just due to dynamical timescales becoming shorter. In addition, the authors argue that the shape of the accretion zone changes outside vs inside of P=60d. I agree that there is a qualitative change, but to me, this change happens only at 10-20 orbits. Lastly, it seems to me that the value for the boundary could be quite sensitive to the random seed. If for example in the fdHi simulation the body at ~ 55d had a tilde b of less than 2sqrt(3), then one could immediately argue that that P~40d-50d is a more appropriate value for the location of the accretion mode transition.
In summary, to me it seems like the presented simulations are much more indicative of the mode of accretion somewhat smoothly changing from oligarchic to 'high-alpha' as alpha is increased, and that it is therefore challenging defining a clear boundary between the two modes.
I suggest that the authors either soften their statements on this, or present additional, quantitative arguments supporting their claims about the universality of this sharp boundary.

	Update figure 6 to show the time evolution in units of the local collision timescale. After doing so, the oligarchic curves lie on top of each other, and the inner non-oligarchic ones don't. This implies two qualitatively different accretion modes.

	Update figure 8 to show multiple versions of each simulation. Gives a much clearer estimate of the error bars on the P=60d estimate.

2) Related to my first point, I think a very interesting, yet somewhat unanswered question is what alpha exactly corresponds to the transition between accretion modes. The authors allude to alpha = 0.1 being of importance, but this choice currently feels somewhat arbitrary. For example, in Figure 11, the authors show the alpha profile for their simulations and show that alpha falling below 0.1 approximately corresponds to tilde b of the largest bodies falling below the critical value of 2sqrt(3). Is there a specific reason for this? Or is this what motivates the choice of alpha = 0.1 in the first place? If so, I think this is an interesting finding, because it allows for a direct prediction of the dominant mode of accretion given the size and location of a planetesimal. Also, how consistent is this alpha = 0.1 value with the simulations with different surface density power laws listed in Table 2?
I recommend that this, in my opinion, quite interesting point is discussed in more detail in the manuscript.
The authors utilize a simplified collision model in which each collision results in a merger. They also perform a test with a merge-bounce model in which some collisions result in an inelastic bounce rather than a merger. They argue that physical collisions are somewhere in-between, and, given that there is little qualitative difference between only-mergers and bounce-merger models, the model outcome should well represent the outcome of physical collisions.

	alpha = 0.1 entirely comes from looking at figure 11. It is simply the location in the disk where the feeding zones settle down to 2*sqrt(3). alpha < 0.1 roughly corresponds to "alpha << 1". Show the b_tilde plots alongsize the alpha curve for the other simulations.

3) I am unconvinced that the test of the collision model indeed is an extreme case. In particular, I'm imagining a very destructive, inelastic collision such that all fragments are smaller than the original bodies. Say the fragments follow a power law distribution assuming a collisional cascade (see Dohnanyi 1969, Zvyagina et al 1974). It's not obvious to me that resupplying smaller bodies to the planetesimal population this way would not impact the merger history. In the extreme case, the entire embryo population would be grinded down by destructive collisions. The authors' bounce-merge model, is essentially equivalent to less efficient, but ultimately constructive planetesimal accretion (thus leading to a prolonged runaway growth phase as shown in Fig. 13). If smaller collisional fragments are re-supplied this could have for example consequences for orbital repulsion and gravitational scattering, which as the authors argue in Sect. 5 requires the presence of smaller planetesimals.
Could the authors comment on this? Maybe the authors can argue why collisions are not expected to be destructive, and clarify why resupplying smaller bodies does not affect the subsequent evolution of the system.

	Accretion mode is size independent (only depends on density and orbital period), so the growth for smaller bodies should behave exactly the same. No chance of enhanced orbital repulsion or scattering with smaller bodies. The other concern to adress would be whether material gets ground down and removed faster than the embryos can grow. Need to compare the growth timescale and calculate some sort of mass removal timescale (look into this).

	Mass removal mechanisms: grinding to meter sizes, radial drift
							 grinding to cm size, pr drag
							 grinding to micron size, dust blowout

4) I think the introduction of several physical concepts warrants additional clarification for accessibility and clarity of the manuscript, and sometimes also lack concrete definitions. In particular, this includes isolation mass, the concepts of tidal stability and Hill radius, the eccentricity dispersion e_h, and the exact definition of tilde b. See my specific comments below.

Where should I put these?

5) I have some suggestions for improving the presentation of the results in Figures, to make the science result more clear and intuitively obtainable. In my specific comments below, I included suggestions for the figures.

Specific comments, notes and suggestions.

TODO: First few edits need to be highlighted in manuscript with strikethroughs and bold face.

Abstract
7: Are short period planets that exotic in the broader context of exoplanets?
	Hard to say! We have replaced 'exotic' with 'wide variety'.
14: "[...] population OF dynamically [...]"
	Fixed
17-18: something is missing in the second part of this sentence.
	Missing words at the end of L17: 'investigate *the effects*'

1.Introduction
26: How large are these solids? I assume this is referencing micron sized dust-grains?
	Correct, some text was added to clarify this.
68: Add reference
	Added a handful of references arguing that the asteroids, KBOs and Oort cloud bodies are leftover planetesimals.
	See Duncan, Tremaine Quinn papers (particularly 1988 one + references?)
72: Maybe elaborate on this. How long is the period of quiescence? How does this instability work?
	We intentionally left this part vague as someone really needs to look into this more! There are three relevant timescales here:
	1. The timescale for the instability itself (depends on the orbital properties of the embryos, see Chambers 98)
	2. Dynamical friction between protoplanets and planetesimals. This will shut off once the planetesimals are either scattered away, or slowly accreted onto the embryos. Not sure what the timescale for this would be. Note that this shouldn't operate at short period according to this work!
	3. Tidal eccentricity damping of the embryos due to the gas disk. Should shut off once the gas disk disperses on ~Myr timescales (Beckwith + Sargent 1996).
	TODO: decide what to add to the text here.
88: Add reference maybe.
	I don't know of any references which have tested the applicability of statistical methods during the oligarchic growth phase, but maybe some clarification would help. The runaway growth phase produces a small number of large bodies, which eventually interact and terminate runaway growth (Ida + Makino 1993, Kokubo + Ida 1995). Given how few oligarchs are produced, one would not expect particle-in-a-box methods to be a good approach here. We've clarified this in the text as well.
99: Capitalize "Systems of Tightly-packed Inner Planets"
	Fixed
104: Gas-Disk driven migration → reference?
	Added reference to Izidoro 17 and 21, which invokes gas disk driven migration, followed by resonant trapping and an possible orbital instability to explain the observed architectures of STIPs
116: What is the "isolation mass" initial condition? Maybe isolation mass can be defined
	We will define isolation mass later in the paper. For now, we removed this term and replaced it with 'fully-formed, evenly spaced protoplanets'.
118: I'm not sure if this belongs in the introduction. I think the mention of future work at the end of the discussion is sufficient.
	Agreed. Removed the last sentence of this paragraph.

2. Overview of planetesimal accretion
2.1 Oligarchic and Runaway Growth
144 onwards: indices in math upright
	Fixed
152: Define rms velocity
	TODO: not sure what the referee wants? Its defined in equation 1?
161: What does "orderly" fashion mean exactly?
	This means the mass ratios between bodies tend toward unity. This has been clarified in the text.
174: Add reference.
	Added reference to Lissauer 1987. Also mentioned that pebble accretion is a viable alternative.
198: Ln Lambda approx 10 missing could use a reference.
	Typo in the text, Ln Lambda ~ 3 is typically used. Added references to Ida 1990, Stewart and Ida 1993). Ln Lambda = 3 was used to construct figure 1, so no changes to the figure needed.
204: I suggest more formally introducing the concept of Hill sphere and hill factor.
	TODO: Where to put this?
209: This sentence seems not necessary.
	Agreed, removed
213: add "dynamically" heating
	Added
222: add "a" crucial assumption
	Fixed
226: Maybe clarify that growth timescale equals collision time scale.
	We have replaced 'growth' with 'collision' here, and done the same elsewhere in the paper for consistency
238: The eccentricity dispersion e_h is never defined anywhere. It is unclear to me what scaling e_h with the Hill-factor means physically.
	e_h can be thought of as the radial excursion of the body's orbit in units of its Hill radius. For e_h > 1, the random velocity of a body dominates over the velocity difference due to the shearing of the disk (over a separation of 1 r_h) and encounters between bodies can be treated with a two-body formalism. We have decided to remove any mention of e_h here, and instead save this definition for section 4 (TODO: define in section 4), where it is next mentioned.
Figure .1:
Since this is an analytical plot, I think one could color the entire parameter space according to the ratio t_relax/t_coll. Currently, it may suggest that the plotted values for e_h are chosen for a particular reason.
	Get rid of the legend, explain that e_h values increase from bottom to top.
Also, I think it would be nice to have a reference for what values of e_h are standard for certain disk conditions, i.e. how outrageous would it be to have a e_h = 16? Are all disks like that? Or do we expect e_h = 1 as the standard?
	Not sure if there is going to be a reference for this. Equilibrium e_h depends on particle mass and gas density. Could try to plot things in terms of that instead. (Maybe fix rho_gas and vary mass?)
Also, I recommend adjusting the x-limits of the plot to correspond to the limits of the orbital period.
	Done

2.2 Planetesimal Size and Extent of Hill Sphere
253: "will remain proportional" - Proportional to what?
	Changed the wording to 'ratio between the relaxation and collision timescale remains constant' to clarify this.
254: but this doesn't have to be the case at all orbital periods, right? Given Fig.1 it seems clear that outside of 1AU, and especially for the Kuiper Belt this condition seems to be fulfilled.
	Exactly! That's the point we're getting at here. (TODO: do I need to clarify something in the paper here?)
263: Again, I think the concept of Hill radius and tidal stability needs to be formally introduced earlier.
	TODO: Introduce this formally earlier
273: Wouldn't these bodies then be tidally disrupted, and therefore not exist? Alpha > 1 is certainly not true for any solar system bodies that I am aware of. I understand that the scaling of alpha predicts alpha >1 at some radius (which I would explicitly mention), but this just implies that inside this radius, bodies must have greater densities to not be disrupted.
	The distance at which alpha = 1 (a_alpha) and the Roche (a_roche) limit are similar, but not identical. For a rigid spherical body, a_roche ~ 0.88 a_alpha. Applying the hydrostatic correction, a_roche ~ 0.6 a_alpha. This is important, because it implies that the new growth mode presented in this paper should be relevant for ring systems! We will add a note of this in the text. I think some of the confusion comes from the fact that the Roche limit and the Hill sphere size are not derived using the same formalism.
289: gravitational radius: is this physically motivated? Maybe reference
	There is a physical explanation of r_g immediately below equation 5 in the text. Also added a reference to Ida 1990.
296: beta > 1 means collision cross section > geometric and small gravitational focusing. I recommend explicitly stating what the physical intuition for beta < 1 is.
	Done.

I think this section could be improved by having a schematic of the three sizes, for two different sets of alpha and beta, that highlights the physical meaning of both. This would communicate intuition for the two dimensionless parameters, and thus allow for a smoother read-through of the N-body results.

3. Numerical Methods
335: children: Would those not be the parent particles?
	We've been representing the collision history of a particle with a tree structure. Following computer science convention, our tree is flipped upside down, with the original planetesimals at the bottom (hence calling them children). I agree that calling them 'parents' is probably more intuitive for most readers, however. We have changed 'children' to 'parents' here.

4. Narrow Annulus Simulations
352: I think it's somewhat ambitious calling four simulations a parameter study.
	Agreed. This has been changed to 'an exploration of the (alpha, beta) parameter space'.
362: Regarding the initial condition, I recommend adding some more intuitive mass units, i.e. approx 0.5 Ceres masses, as well as the corresponding radius 341 km. Also, to what is the bulk density reduced to? Is this still physical? What is the corresponding radius? Maybe motivate the initial planetesimal masses / radii (e.g., Morbidelli 2009, Hartlep & Cuzzi 2020, Gerbig & Li, 2023). Also, maybe put the total mass of the planetesimals into context. I think this is 0.33 Earth masses, within 0.1 AU of Earth, the feeding zone of Earth.
	I don't think I like the idea of providing more physical motivation here. This was meant to be a study of the effects of varying alpha and beta. I think some of these points should be saved for the full disk section. Perhaps it would be better if I described everything in dimensionless units? TODO

367: I'm assuming the semi-major axis is also drawn from a uniform distribution?
	Yes, this is now clarified.

377: is alpha=1 chosen because it is the highest stable alpha? If so I would write that. Why is beta = 15000 chosen?
	Alpha = 1 corresponds to r_pl = r_hill. One would expect that gravitational scattering should completely shut off for alpha >= 1. Beta = 15,000 corresponds to <e^{1/2}>^2 = 0.05, which is the largest eccentricity for which a particle at 1 AU will an apo and pericenter distance that is still within the boundaries of the disk. For an eccentricity dispersion this large, the stirring timescale is much longer than the simulation time, so we can be confident that the boundary effects do not get worse. An explanation of the large beta value has been added to the text.

Figure.2: Some suggestions to improve this figure are:
A reference planetesimal size that shows what mass a circle has
	Might skip this, want to keep everything in dimensionless units here TODO: decide
Horizontal lines in each panel showing where e_h is
	Might skip this, depending on whether I remove the e_h lines from figure 1. TODO
Maybe consider having beta value as a column label and alpha value as a row label. I would also explicitly state the values for alpha.

I would also probably label the four panels / cases a,b,c,d and then refer as such in the text body
	Done

391: Why would this because there are no high eccentricity planetesimals. Isn't the fact that there is no residual population not already evidence that encounters are accreting rather than scattering? Also, isn't this also not just very expected since alpha exactly quantifies the relative importance of scattering vs accretion?
	I'm assuming this is referring to the statement that begins with 'The lack of high eccentricity planetesimals...'. I think you must be misinterpreting what we said in the text here, because it sounds like we are both making the same point. When we turn up alpha, the high eccentricity planetesimals dissapear. This is expected, because large alpha makes gravitational scattering difficult, so those bodies that undergo close encounters end up being accreted by the protoplanets, rather than getting kicked onto high eccentricity orbits. If there is something that needs to be clarified in the text, please let us know!

394: small alpha, large beta doesn't look dynamically cool to me.
	This should have been referring to the dynamical state of the growing protoplanets, relative to the initial eccentricities of the planetesimals. This has now been clarified in the text.
400: Small alpha large beta does not look dynamically cooler than large alpha large beta to me. In fact, the opposite.
	You are correct! This is a mistake in the text (replaced 'latter' with 'former').
402: I don't really understand why small alpha small beta leads to so many dynamically hot bodies, both planetesimals and protoplanets. Maybe the authors can elaborate.
	When alpha and beta are both small, the gravitational stirring timescale is very short and the disk quickly heats up. Because the stirring timescale goes as v^{1/4} (TODO: reference), the scattering slows down as the disk gets hotter. So runaway growth doesn't really kick in until after the eccentricities have grown enough that stirring slows down. I think this is why the initial choice of beta doesn't really seem to affect the final dynamical state of the disk in figure 2. This is also why the initial choice of e_h isn't very consistent across planetesimal accretion studies (cite some examples). TODO: Let's add another paragraph discussing this here.

Figure 3:
Again, I really think labeling the cases would be helpful.
I would express the mass in units of e.g. Ceres
	Todo: Push back on this? Use dimensionless units for this section.
406: I'm not sure about "significantly" larger. The biggest body in large alpha, small beta is about a factor 3 more massive than in small alpha, small beta. This is only a factor of 1.4 in size.
	The text now says 'a factor of a few larger'
415: the y-label in Fig.6 does not match the description in the text
	This comment must be referring to figure 4? If so, 'm_max' in the text has been changed to 'M'.
Figure 4:
Why don't the simulations start all with m_max/m = 1?
I'd be interested in also having another panel that shows the evolution of just M_max and <m> (maybe in solid and dashed lines) for the four simulations. That would also underscore that the turnover is driven by an increase in <m> rather than m_max.
	We have now added a second panel to figure 4 which shows the evolution of these quantities separately. Todo: Units? Why does the max mass decrease for case a?

418: I would specifically state that for the large alpha cases runaway growth only occurs at the start, to avoid initial confusion that the slope turns around when the planetesimal supply is depleted.
435: Is this true? I was under the impression that the residual planetesimal population remains? I think this does not seem consistent with the rest of the work...?
	TODO: Change the wording here. Maybe this isnt true. Would the growth mode simply 'flatten' out here?
446: What exactly does alpha being non-negligible mean? Given that alpha is always less than unity, what alpha value marks the transition of the oligarchic growth regime vs, consumption regime? See major comment.
	This is a good point, and I'm not sure if we have a definitive answer. Inspecting figure 11, theh transition seems to occur near alpha ~ 0.1. I think the best thing to do here is to state the converse: for alpha << 1, most encounters result in scattering and you get oligarchic growth. The key finding from this section is that oligarchic growth does not occur when alpha is not << 1. The text has been updated to reflect this.

5. Full disk Simulation
5.1 Initial Conditions
470: Reference of TRAPPIST-1
	Added reference to Gillon 2016 + 2017, Agol 2021
483 I recommend writing Eq. (6) along the lines of 10 g/cm^2 \times A (M_*/M_\odot)(r/1AU)^(-delta) for readability. The Sigma_MMSN is never used so it's not necessary.
	Done
Table 1.
I don't think the notes are necessary
I would probably put the planetesimal masses in units of Ceres
	Where should I put the units?
I think Table 1 and Table 2 could be combined
	Table 1 has been merged into table 2.

490: How could such an enhancement in planetesimal surface density be achieved initially? Is this initial condition consistent with planetesimal formation models, e.g. Dra̧żkowska & Dullemond 2018, Lenz et al 2019.
	This is briefly discussed in the introduction but we will expand upon it here. Most mechanisms involve dust traps or pressure bumps occuring at some preferential location in the disk which trigger the streaming instability. Dra̧żkowska & Dullemond 2018 show that a migrating snow line can actually trigger planetesimal formation over a fairly wide area, and also produce surface density enhancements at least as large as the ones we use. Given that low-mass stars have particularly extreme luminosity changes during the pre-main sequence (see Baraffe 2015), one can imagine that it might not be hard to trigger planetesimal formation over a wide area. This particular paragraph in the text has been expanded to discuss this.

493: What is the motivation behind variation in the power-law index? Is there significant uncertainty in the MMSN power law index?
	Chiang and Laughlin 2013 applied a similar analysis to Kepler multis and found that delta varies between 1.6 and 1.8. Sub-mm observations of cold outer disks finds power law slopes as low as 0.5 (Mundy 2000, Andrews 2009, 2010). Therefore, we vary delta by +/- 1.0 from the MMSN value. There is now a separate paragraph in the text explaining this
499: Maybe elaborate on why this equilibrium is physical? Are there other works which use the same balance to inform initial conditions?
	t_vs = t_gas is more of an upper limit on the eccentricity, although this is the same technique that Kokubo + Ida 1998 use to set up their initial conditions. It's certainly possible that the disk could have started out dynamically colder, although as we showed in figure 2, this doesn't seem to affect the resulting protoplanets. We have added a short discussion of these points to the text.
504: missing respectively. And semi-major axes.
	Fixed
505: When is the Stokes regime valid, and what does it mean?
	In the Stokes regime, the mean free path of the gas is much smaller than the size of the bodies and the gas can be treated as a fluid. For conditions in a typical protoplanetary disk, Stokes drag is relevant for cm-sized bodies and larger. This has been added to the text.
506: The letter tau is used for a timescale just above, but now it says t. I would recommend staying consistent. Tau tends to be used for dimensionless timescales, so I would probably go with t_s and t_vs.
	tau_vs has been changed to t_vs in equation 7.
511: Solid-to-gas surface density ratio of 1/240 seems very small, especially given that the authors are considering the solid surface density to already be enhanced by a factor of 100. Also, if the planetesimals are already formed, it is conceivable that the gas disk is already dissipating (especially so close in). I think the authors could be over-estimating the gas density by up to a factor of a 100, which means the stopping time would be a factor of 100 longer.
	TODO: Fix
516: I suggest using another symbol for the temperature power law, since Q is associated with the Toomre value of the disk, for example q. Also, this equation should read T(r) = T_0(r/1AU)^-q, right?
	Q was changed to q to be consistent with the Andrews + Williams paper. the '/1AU' has been added to the T(r) equation.
521: Where is that equation coming from? Maybe add reference.
	This is equation 4.30 from Phil Armitage's 'Astrophysics of Planet Formation'. A reference has been added to the text.

5.2. Gas Drag Force
512: I think it would be very informative to calculate the dimensionless stopping time, i.e. the Stokes number of the planetesimals, I'm assuming that they are substantially above unity, which means that the bodies are essentially decoupled from the gas.
	Agreed. Describing things in terms of the Stokes number allows us to simplify this section a bit. For the initial planetesimals, St >> 1 at both the inner and outer disk edge Because St ~ m^1/3, we can guarantee that bodies will remain decoupled from the gas throughout the simulation. By discussing the Stokes number, we can also avoid having to compare the stopping time with the integration time of the simulations. These points are now all discussed in the text.
536: Isolation mass lacks definition
	We have removed any mention of the isolation mass here. The fact that St ~ m^1/3 is enough to guarantee that the gas and solids remain decoupled.

5.3 Time-stepping Criterion
554: footnote 2 can just go in the main text body.
	Done
584: footnote 3 can also go in the main text body.
	Done
584: Are there any works on the numerical robustness of this more efficient time-stepping scheme. While I'm inclined to believe that the error is only marginal, I think having some sort of test (or a reference to one) would make the results more steadfast.
	I'm fairly certain that we are the first ones to try this. I've added an appendix that compares this timestepping scheme against a fixed timestep case. There appear to be no meaningful differences. TODO: Show mass distribution as differential or cumulative?

5.4 Results
586: Maybe include reference to figure where this is shown
	Figure 1 of Kokbubo 2002 shows this clearly. A reference to the paper has been added in the text.
592: Is this really always true? Counterexample e.g. Emsenhuber et al 2021 and others.
	The NGPPS models (along with the Clement 2020 paper) seem to be a couple of notable exceptions to this, although the embryo growth is not treated self-consistently leading up to the giant impact phase. We have added a mention of the Emsenhuber papers to the text.
Table 2:
What is the rescaled integration time?
	TODO: Not sure where to explain table columns now?

Figure 5:
I think the bottom panel should have a log y-axis
I would make the color of the dots consistent with Fig.2, where the final outcome is orange
There could be numbers on the contours indicating the density of the initial condition.
The meaning of the dashed line in the bottom panel could be indicated inside the figure itself.
I'm confused by the statement "interior to the oligarchic growth region (around 60d)". Where does this definition come from? Is this defined by bodies exceeding the critical feeding size? If so the statement should be rephrased to make the causal connection clear.
	The y axis on the bottom panel of figure 5 is now log-scaled. Point colors in both panels have been changed to orange, to be consistent with figure 2. The dashed line in the bottom panel is now labeled. Yes, we are defining the oligarchic growth region to be the region where \tilde{b}_required <= 2\sqrt{3}. The last sentence in the figure caption has been rephrased to make this clearer.
606: How can a mass be expressed in units of feeding zone size?
	We are solving equation 11 for \tilde{b} and using the initial local surface density, along with the final particle mass to calculate the feeding zone size. We tried to explain this starting on line 611, but it sounds like some further clarification is needed. Some additional text has been added to the paragraph following equation 11 to explain where the value for \tilde{b} comes from.
620: What is the intuition behind tilde b, how exactly is it defined? Since it is a dimensionless quantity, calling it an orbital separation is not intuitive to me. Can tilde b be converted to a physical size? Also, I struggle to interpret the critical tilde be of 2*sqrt(3). Do I understand correctly that a body with tilde b less than 2 sqrt(3) is associated with a feeding zone so small that close encounters are not possible, and thus are not expected to grow further?
	This is buried in some of the references in this section (Nazakawa + Ida 1988, Kokubo 1998, Kokubo 2002), but I'm realizing this is not stated explicitly anywhere in the paper so I will do so: \tilde{b} is a length in units of the local Hill radius. 
	The \tilde{b} = 2\sqrt{3} value is important because this is the smallest possible feeding zone size that a body can have. As eccentricity decreases, the feeding zone size decreases, but has a minimum value of 2\sqrt{3} for (e,i) = (0,0) (See section 6 of Nazakawa + Ida 1988 for an explanation as to why). Therefore, a protoplanet undergoing oligarchic growth (in which gravitational scattering events are frequent and keeps the largest bodies on circular orbits) will not accrete more than 2\sqrt{3}*r_h*Sigma worth of material from its surroundings. The key takeaway from the bottom panel of figure 5 is that there are bodies interior to 60d that must have accreted material from a much wider annulus than this, and so oligarchic growth is not operating here.
	TODO: Update text somewhere?
632: Why are the bodies expected to follow the tilde b = 2sqrt(3) from a theoretical perspective? Also, I'm unconvinced that the bodies do that at all in the presented simulations - they fall below this value but don't follow it.
	See response to 620. You are right that the outer disk bodies do not all reach \tilde{b} = 2\sqrt{3}. TODO: Should probably mention that this should be interpreted as a maximum. Not feasible to run the outer disk to completetion (especially in the fdSteep case).
640: If anything the opposite seems to be true, no? Maybe have a specific value for the expected eccentricity of protoplanets with and without inelastic damping.
	TODO
644: If I look at the contour map by eye, the eccentricity distribution doesn't seem all that different in the inner vs outer disk. Maybe the difference can be quantified?
	You are right that the variation in eccentricity is not huge. For the model used by our initial condition generator, the mean eccentricity is a factor of 4x larger at the outer disk edge, compared to the inner disk edge. This is now stated in the text.
647: This is an interesting point. I'd be interested in seeing a comparison between inner and outer disk accounting for that difference in time scales. Is it possible to include such a comparison maybe with a figure?
	Yes! See our updated version of figure 6. The time is now shown in units of the local accretion timescale, and it is now apparent that the planetesimal surface density curves all follow the same trajectory exterior of 60d.
Figure 6:
This figure could be improved if each line was colored by the semi-major axis of the radius bin, and if there was a corresponding colorbar. The difference between the 60 day boundary could still be indicated with an additional line.
	See our response to the previous concern, along with our response to major comment #1.
	TODO: Update text to reflect changes to figure 6.
	TODO: Run fdhi for a thousand steps at higher output cadence, so the innermost curve starts sooner
665: I struggle to see that there is a significant qualitative difference in the behavior in the lines just inside and just outside the boundary of 60 orbits. The first dashed line also seems to converge to a sigma/Sigma \sim 5 %, not much below the first solid line.
	See our response to the previous concern, along with our response to major comment #1.
668: add reference. Also footnote 5 can probably go in the main text. Can the authors provide an equation that converts tilde b to units of r_h?
	\tilde{b} is already in units of r_h. This is now clarified near line 620 in the text. The 30 percent value simply comes from removing the size of the feeding zone (2\sqrt{3}) from the spacing between oligarchs (10). This is now clarified in the text here. I think the footnote text can be removed after adding this clarification.
682: This statement seems only to be true for fdshallow. The others have a spread in eccentricity that covers two orders of magnitude...
	I think we need to change the wording here. What we were trying to say is: the resulting a-e distributions for the embryos between all of the simulations are quite similar (with the exception of fdLo), despite the fact that the a-e distributions of the initial conditions are different.
685: It is not clear to me why more inelastic collisions lead to a more sharply peaked eccentricity distribution, and not just more damping like the authors argue for the fdLo simulation.
	I'm not sure what you mean by 'more shaply peaked'? The argument we're trying to make here is this: In the case that the initial surface density is large (e.g. the inner part of the fdSteep simulation), the initial eccentricity dispersion is larger (because the viscous stirring timescale is shorter and so we start out with a larger <e^2>^1/2 to make t_vs = t_gas). The increased surface density leads to more collisions, which leads to more inelastic damping. So the larger eccentricity dispersion that we started with quickly damps out.
701: Again, I don't see that the value for tilde b approaches 2sqrt(3) beyond P=60d orbits. The bodies have smaller feeding zones than tilde b = 2sqrt(3)
	Perhaps if we ran some of these simulations for longer, the embryos beyond 60d would all reach \tilde{b} = 2\sqrt{3}. But this requires an intractible number of timesteps, particularly for the fdSteep simulation, where the surface density in the outer disk is quite low. I think the best we can say here is that none of the embryos exceed \tilde{b} = 2\sqrt{3} which is consistent with the oligarchic growth model.
702: Are they that similar? Maybe this statement can be more quantitative? To me it looks like the values fall below 2sqrt(3) at 40 days for fdHiShallow and at 70 days for fdLo. The interpretation that the boundary is only driven by the local value for alpha seems somewhat speculative. I'd also be interested how randomness affects this, i.e. if the simulation is initialized with a different random seed, by how much does the final distribution change?
	We were saving these for the follow up paper, but we do have multiple versions of fdHi, fdSteep and fdShallow! These have now been added to figure 8. The added data does make it clear that the boundary is quite fuzzy. I've also plotted the mean and standard deviation of the \tilde{b} in a number of orbital period bins. I think this makes the accretion boundary a little more apparent. With the exception of maybe the fdShallow simulation, the locations in the disk where \tilde{b} drops below 2\sqrt{3} are broadly consistent, which supports our claim that the boundary location is driven by the local value of alpha.
	TODO: Update text to recflect changes to figure.

5.5. Assembly History of Embryos
719: add references.
	Done
721: Does this distinction have analogies in galaxy formation? Also isn't this overly strict? I would expect that the initial population has quickly collided at least once, so most of the collisions will be major mergers under this definition. This seems to be supported by Fig. 9 where the smooth accretion fraction is always <10%.
	TODO: Ask tom about this?
724: Why are only the 100 largest bodies chosen? It seems more natural to me to choose a mass cut-off. Also, given that in Fig9 is plotting mass on a log-scale, I suggest just plotting the entire population including those planetesimals that remained relatively small and pristine. Also, why is the fdHi simulation chosen?
	TODO: What else to include here? Everything?
726: I think it could be explicitly stated that smooth accretion = minor mergers.
730 & onwards: Can the authors quantify this trend? E.g., by making three linear fits (with error bars) for > 60d, < 60d and the full population. I'm in particular unconvinced that the trend for P > 60 days truly is flat, given the small sample size for larger bodies.
Figure 9:
Given that I'm skeptical about the 60 day boundary being clear and universal, I think this plot would be stronger if the bodies were colored by orbital period (in addition to the marker style indicating this boundary), and the plot would include a corresponding colorbar.
The authors state "there is a decreasing correlation between smooth accretion fraction and mass, which suggests that most growth occurs between equal mass bodies". As far as I understand, this doesn't have to be the case. I would argue that more massive bodies require more time to build up and during later stages there are simply less primordial planetesimals left so collisions are just more likely to be major mergers.
The authors state "indicating that accretion of planetesimals is important during all phases of the evolution", yet the smooth accreting fraction is only between 1-10% for all these bodies. Seems to me that for all bodies, regardless of mass and orbital period, major mergers are significantly more important - that is at least with the author's definition of major and minor merger.
	I agree that 1-10% is a small fraction. The mass cutoff should definitely be larger. TODO: Change mass cutoff and hopefully the same scaling with mass remains.
In the legend, I think "orbit" should be put in the subscript instead of P_orbit
746: This is just because in the inner disk there are no primordial planetesimals left to build the more massive bodies, right?
	No, the accretion mode should be qualitatively different, even before the planetesimals run out. TODO: Hopefully including more data in the plot and maybe increasing the mass cutoff for 'smooth' accretion will make this more obvious.
750: use /citep[see][]{Kokubo}
	Fixed
756: I agree that the large bodies must've originated due to mergers with other embryos, but this is not really an explanation for why the feeding zone size doesn't match the theoretical value, no? Instead it's the lack of gravitational scattering, as the authors write above. Maybe the authors can clarify specifically why oligarchic growth models predict this feeding zone size, and also why gravitational scattering is suppressed in the inner disk (due to the lack of primordial planetesimals in the inner disk?).
766: "One should note that [...]." This sentence can be deleted in my opinion.
	Removed
Figure 10:
I suggest indicating the boundary between accretion modes in this figure. >60d orbits only include 4-5 bodies. However, to me it seems like the shape of the accretion zone is qualitatively the same until ~ 20 orbits, which strengthens my view that the 60d orbit boundary is not very clear at all.
I suggest matching the x-axis limits the shown xdata.
I think "Fraction of planetesimals accreted" is a somewhat misleading description of what the plot is showing. What about: "Distribution of accreted planetesimals"?
775: The transition between smooth and unimodal seems to be happening at about 20 days to me.
Figure 11
The y-axis label overlap
I think the finding that the crossing of alpha = 0.1 in the bottom panel well matches the crossing of tilde b = 2 sqrt(3) is interesting and worth highlighting more, maybe with some vertical lines? Is there a meaning behind alpha = 0.1, i.e., is this coming from some analytical consideration? See my major comments

6. Simplifying assumptions
6.1 Collision cross section
802: As I argued above, I'm unconvinced in this clear-cut boundary at 60 days. Maybe the authors could put a quantitative uncertainty to this number?
816: I'm not following where the factor of 15 comes from. Maybe the authors can write down the explicit proportionality used to obtain this value.
821: How is this expected to change with f?
843: Maybe add reference to such short period planets.
868: I'm not sure I fully follow this statement of the collision model. See my major comment above.

Figure 12:
Add a description of what green and orange means to the plot.

901: Are the populations really indistinguishable? The authors could perform for example a Kolmogorov-Smirnov test to underline this statement. If not, I would suggest softening this statement.
912: The authors perform this collisional model test very close-in where residual planetesimal populations are suppressed anyways. Maybe the authors can comment on what would change (if anything) if this test was performed in the outer parts of the disks.

7. Summary and Discussion
933: I would think the gravitational feedback still operates, just on a timescale that is much longer than the planetesimal accretion timescale that depletes the planetesimal supply.
942: Refer to the section where this is discussed.
946: alpha is already reviewed in the paragraph above
952: non-negligible is not very precise . Maybe change to while "alpha close to unity"
954: Refer to figure where this is shown. I think Fig.2? Also "the same" is very strong. Maybe "very similar" instead.
959: Sentence is currently implying that the boundary location is set by the distance to the star which doesn't really work in my opinion. Maybe just delete "and therefore the location of the boundary between these accretion modes" as this becomes clear with the subsequent sentence anyways.
965: I think this is too strong a statement. See my previous comments. I think the subsequent sentence "Although [...]" is more appropriate.
972: Refer to Fig. 10.
978: Missing a verb, maybe "differs" ? Also the second part of the sentence "along with [...]" seems odd to me. Maybe rephrase it in two sentences?
986: Is there a reference for this? It does not seem obvious to me. See my previous comments.
996: what bulk density? Maybe also rephrase to a less strong statement, e.g., "[...] bulk density, our results suggest that the boundary is located around 5 days in orbital period."
998: I recommend introducing a subsection heading here, for clarity.
999: Maybe add references of what values for alpha typically were used?
1000: This sentence is somewhat odd to me. Yes, alpha = 1 are the largest bodies stable under tidal gravity, however, this applies not just to circumplanetary disks but also for the circumstellar planetesimals the authors investigated in their work. Because of this, it doesn't seem like this statement is an appropriate segue to circumplanetary disks. I suggest separating this paragraph as a subsection under a heading such as "high-alpha accretion in circumplanetary disks".
1003: delete "small"
1007: Can the authors clarify the relevance of the inclusion of spiral density wave interactions?
1018: I recommend defining R_U for readers unfamiliar with Ida et al 2020, e.g., by summarizing that Ida et al 2020 considered a circumplanetary disk around Uranus.